---
title: "p8105_hw1_bz2570"
author: "Boran Zhai"
date: "2025-09-18"
output: github_document
---
## Problem 1
#### Load the library and the data
```{r}
library(moderndive)
library(tidyverse)
library(ggplot2)
```
```{r}
data("early_january_weather")
```

#### Write a short description of the dataset
```{r}
vars <- names(early_january_weather)
n_rows = nrow(early_january_weather)
n_cols = ncol(early_january_weather)
temp_mean = mean(early_january_weather$temp)
summary(early_january_weather)
cat(" names:",vars,"\n",
    "size:",n_rows,"rows","and",n_cols,"columns","\n",
    "the mean temperature:", temp_mean)
```
##### Description 
The dataset early_january_weather contains 358 rows and 15 columns, with the following variables and some values of important variables: origin, year(2013), month(January), day(from 1-15), hour, temp(ranges from 24.08 to 57.92 with a mean of about 39.58), dewp(ranges from 8.96 to 53.06 with a mean of about 28.06), humid(ranges from 32.86 to 100.00 with a mean of about 65.48), wind_dir, wind_speed, wind_gust, precip, pressure(ranges from 1011 to 1034 with mean of about 1023), visib(ranges from 0.12 to 10.00 with mean near 8.52), and time_hour(falls between 2013-01-01 01:00:00 to 2013-01-15 23:00:00). <br/>
  
The mean temperature is approximately 39.58.

#### Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable 
```{r}
Sys.setlocale("LC_TIME","English")
scatterplot <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(size = 2, alpha = 0.8) +        
  scale_color_gradient(low = "blue", high = "red") + 
  labs(x = "Time_hour", y = "Temperature", color = "Humidity") +
  theme_minimal()
print(scatterplot)

ggsave(
  filename = "D:/zbr/p8105_hw1_bz2570/scatterplot.png", 
  plot = scatterplot
)
```
##### Description
The scatter plot illustrates the relationship between temperature(y-axis) and time_hour(x-axis) for the early January.It shows a moderate overall rise of temperature from January 1st to 15th, with daily minimum and maximum temperatures gradually increasing. This pattern was temporarily broken by two anomalous periods around January 7th-8th and January 15th. Besides, daily temperatures follow a regular and wave-like pattern. Within each day, lower temperatures usually occur overnight and in the early morning, while higher temperatures are observed during midday and afternoon hours. 
Additionally, each point colored according to the humidity level, using a blue-to-red gradient where blue represents low humidity and red represents high humidity. However, there is an inverse relationship between temperature and humidity: warmer times always have lower humidity(blue) and colder times have higher humidity(red). 

## Problem 2
#### Create a data frame
```{r}
set.seed(917)
df <- data.frame(
  random_sample = rnorm(10),
  logical_vector = rnorm(10) > 0,
  char_vector = letters[1:10], 
  factor_vector = factor(rep(c("Small", "Medium", "Large"), length.out = 10)) 
)
str(df)
```

#### Try to take the mean of each variable and explain
```{r}
mean_random <- mean(pull(df, random_sample))
mean_logical <- mean(pull(df, logical_vector))
mean_char <- mean(pull(df, char_vector))
mean_factor <- mean(pull(df, factor_vector))
cat(" The mean of random sample is",mean_random,"\n",
    "The mean of logical vector is",mean_logical,"\n",
    "The mean of character vector  is",mean_char,"\n",
    "The mean of factor vector is",mean_factor
)
```
